---
header-includes: 
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H}
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "A Case Study on Chemical Flashiness of Streams and Canals in Florida"
subtitle: "https://github.com/ytgong/ENV322_group_project"
author: "Ethan Ready, Gabi Richichi, Theo Cai, Yutao Gong"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage

```{r setup, include=FALSE}
# Set your working directory
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE, figure.cap = TRUE, figure.pos = "H")

library(dataRetrieval)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(knitr)
library(sf)
library(maps)
library(cowplot)
library(gridExtra)
library(grid)
library(float)

#Set theme
theme_set(theme_bw())

# Load your datasets
BLUEdata <- read.csv("../FL_CSV/hyp1/BLUEdata.csv")
BOWdata <- read.csv("../FL_CSV/hyp1/BOWdata.csv")
CALdata <- read.csv("../FL_CSV/hyp1/CALdata.csv")
CHASdata <- read.csv("../FL_CSV/hyp1/CHASdata.csv")
CRANEdata <- read.csv("../FL_CSV/hyp1/CRANEdata.csv")
DRAINdata <- read.csv("../FL_CSV/hyp1/DRAINdata.csv")
FANdata <- read.csv("../FL_CSV/hyp1/FANdata.csv")
FELLdata <- read.csv("../FL_CSV/hyp1/FELLdata.csv")
HOLdata <- read.csv("../FL_CSV/hyp1/HOLdata.csv")
HOMdata <- read.csv("../FL_CSV/hyp1/HOMdata.csv")
HUNdata <- read.csv("../FL_CSV/hyp1/HUNdata.csv")
ICHEdata <- read.csv("../FL_CSV/hyp1/ICHEdata.csv")
MADdata <- read.csv("../FL_CSV/hyp1/MADdata.csv")
MAINdata <- read.csv("../FL_CSV/hyp1/MAINdata.csv")
MANdata <- read.csv("../FL_CSV/hyp1/MANdata.csv")
RAINdata <- read.csv("../FL_CSV/hyp1/RAINdata.csv")
SANTAdata <- read.csv("../FL_CSV/hyp1/SANTAdata.csv")
THREEdata <- read.csv("../FL_CSV/hyp1/THREEdata.csv")
TURKdata <- read.csv("../FL_CSV/hyp1/TURKdata.csv")
TURNdata <- read.csv("../FL_CSV/hyp1/TURNdata.csv")
WACdata <- read.csv("../FL_CSV/hyp1/WACdata.csv")

hyp2.full_cleaned = read.csv("../Processed_Dataframes/hyp2+chemo.csv")
violin_comb = read.csv("../Processed_Dataframes/violin_comb.csv")
```


# Rationale and Research Questions

## Rationale
Storm events have the potential to influence rivers and streams by temporarily altering their discharge levels and chemical composition. Rapid changes in stream discharge is known as hydrologic flashiness (1). A stream that “experiences a rapid increase in flow shortly after onset of a precipitation event, and an equally rapid return to base conditions shortly after the end of the precipitation event” is considered to be hydrologically “flashy” (1). Rapid changes in chemical composition, such as nutrient concentrations or conductance levels, is known as chemical flashiness or a departure from chemostasis. “Chemostasis occurs when DOM [dissolved organic matter] remains unchanged despite changes in discharge” (2). 

Both hydrologic and chemical flashiness can alter the natural processes of rivers and streams, particularly if these rivers and streams are experiencing an increase in flashiness over time. These alterations can disrupt human activity such as hunting, fishing, recreation, and the utilization of water for drinking, and, as a result, people incur costs. One consequence of hydrologic flashiness is flooding of nearby lands. Flooding events in the U.S. cost an average of 4.4 billion USD per event in damages (5). One alteration caused by storm-induced flashiness is eutrophication. Eutrophication “is characterized by excessive plant and algal growth due to the increased availability of one or more limiting growth factors needed for photosynthesis, such as sunlight, carbon dioxide, and nutrient fertilizers” (6). In this case, the limiting growth factor that chemical flashiness can increase is nutrient levels. As nutrient levels increase, this can cause “blooms of blue-green algae(i.e., cyanobacteria), tainted drinking water supplies, degradation of recreational opportunities, and hypoxia. The estimated cost of damage mediated by eutrophication in the U.S. alone is approximately $2.2 billion annually” (7). 

The effects of storm events on bodies of water are compounded by human activity. The use of pesticides and fertilizers on agricultural land, as well as the way in which sewage is managed, can make nutrients such as nitrogen and phosphorus accessible for storm transport in the first place (8). Furthermore, anthropogenically-induced climate change is intensifying natural precipitation events by making them wetter, longer, and more intense (9). As storms worsen, so do their effects. 

This applies to the region of this study in particular. Florida is a state that has a history of extreme weather events such as hurricanes and tropical storms, as well as a history, particularly recently, of eutrophication events. High levels of nutrient loading, in combination with high temperatures that are typical of Florida’s climate, create an environment conducive to eutrophication (5). The Florida Department of Environmental Protection lists more than 1,400 water bodies as “impaired by pollutants” (5). In this study, 21 water bodies including rivers, creeks, springs, and canals in Florida are examined for their hydrologic and chemical flashiness, the urbanness or ruralness of their surrounding lands, as well as the water use levels of the surrounding communities.  

These sites are of great importance to people. For example, Chassahowitzka River is a spring-fed river in a national wildlife refuge, on which people engage in sportfishing, crabbing, camping, and hiking activities (6). The Main Relief Canal in Vero Beach, managed by the Indian River Farms Water Control District, drains much of Indian River County and the City of Vero Beach to prevent flooding on residential and commercial land (10). Turnbull Creek is a “critical waypoint for waters flowing north into Turnbull Bay, Spruce Creek, the Indian River, and Atlantic Ocean” (11) and a habitat for numerous species (12). A majority of the people of New Smyrna Beach (75% of constituents) just voted in favor of a referendum to pay $8.94 million to protect the land along Turnbull Creek from development (13). Many of the sites of this study are well-managed, located in state or national parks, and all of the canals are artificially created. There are already management-schemes and policies in place to protect these water bodies. For example, to reduce nutrient total maximum daily loads, there exists action plans (14) and the Springs Protection Act (15). It is useful, then, to better understand the effects of flashiness on these water bodies in order to inform managers to implement more efficient and cost-effective management practices. 

## Research Questions

The goal of this study was to better understand the effects of storm events on rivers and streams in Florida. The following questions and hypotheses guided the study: 
Question 1: Is there a correlation between hydrologic and chemical flashiness in rivers and streams? 

Hypothesis 1: There is a correlation between hydrologic and chemical flashiness. Sites with high hydrologic flashiness are more likely to be chemically flashy because of shared predictor variables.

Question 2: How do rural / urban features of a site impact its chemical flashiness? 

Hypothesis 2: There is a correlation between the urbanness / ruralness of a site and its chemical flashiness. Sites with high urbanness are more likely to be chemically flashy because of impervious surfaces and non-point source discharge potential.


\newpage

# Dataset Information

Data for 21 water bodies (7 rivers, 7 springs, 3 creeks, and 4 canals) in Florida will be collected. They were chosen by searching all USGS Florida sites with high frequency nitrate and discharge data that had significant periods of overlap. Each site's data was shortened to focus on period of time with consistent data collecting.

For question / hypothesis 1 regarding hydrologic and chemical flashiness, high frequency discharge and nitrate data will be extracted from the USGS NWIS DataRetrieval package. A Richards-Baker Index will be calculated using the discharge data as a measure of hydrologic flashiness. C-Q plots will be generated using the nitrate and discharge data as a measure of chemical flashiness. The data will be combined into one dataframe, along with site information, for trend analysis.

For question / hypothesis 2 regarding the properties of each site that may affect chemical flashiness, data will be retrieved from United States Geological Survey (USGS). Specifically, county population values, irrigation water use levels, thermoelectric water use levels, and a chemostatic coefficient will be examined. Maps taken from ArcGIS will be examined to ascertain the presence or absence of features indicative of urbanness or ruralness surrounding each site, such as airports, commercial center, or farms, as well as the proximity of these features to each water body.


```{r, echo = FALSE, message = FALSE}
df <- data.frame(Variable = c("Nitrate concentrations -- high frequency","Discharge -- high frequency","Richards-Baker Index value", "County population", "Irrigation water use", "Thermoelectric water use", "Chemostatic coefficient"),
                 Units = c("ng/L","m^3/s","", "k people", "Bgal/day", "Bgal/day",""),
                 Type = c("Independent","Independent","Independent","Independent","Independent","Independent","Dependent")
)

kable(df, caption = "Variables used for analysis")

sites.nonalpha<-c("WAC", "MAD", "BLUE", "MAN", "FAN", "SANTA", "ICHE", "HOL",
              "HUN", "HOM", "CHAS", "RAIN", "BOW", "CAL", "THREE", "TURN",
              "DRAIN", "CRANE", "TURK", "FELL", "MAIN")

df2 = read.csv("../Processed_Dataframes/siteinfo.csv") %>%
  select("Site No." = site_no,
         "Station Name" = station_nm,
         "County" = county_nm)%>%
  mutate("Abbr." = sites.nonalpha)

kable(df2, caption = "USGS sites selected for analysis") 
```

\newpage

# Exploratory Analysis 


```{r, include = FALSE}
waterfeatures <- st_read("../other_data/hydrogl020.dbf")
waterfeatures <- filter(waterfeatures, STATE == "FL")
waterfeatures <- filter(waterfeatures, FEATURE != "Apparent Limit" & FEATURE != "Closure Line")
HUC6 <- st_read("../other_data/WBDHU6.dbf")
HUC6.Florida <- HUC6 %>%
  filter(States %in% c("AL,FL", "FL", "FL,GA", "AL,FL,GA"))
waterfeatures <- st_set_crs(waterfeatures, 4269)
states<- st_as_sf(map(database = "state", plot = TRUE, fill = TRUE, col = "white")) 
states.florida = filter(states, ID %in% c("florida"))
hyp2.spatial <- st_as_sf(hyp2.full_cleaned, coords = c("dec_long_va", "dec_lat_va"), crs = 4326)
```

Figure 1 is a map that shows the watershed and water feature information of Florida and location of the sites we chose to analyze. It can be seen that the sites chosen encompass a variety of water features, ranging from streams, banks, to canals. They are also located in different regions of Florida, and the hydrological heterogeneity among sites offers an opportunity to study the potentially different interactions between chemostaticity and hydrological flashiness.

```{r, echo = FALSE, message = FALSE, fig.cap="Location of sites analyzed on watersheds",fig.height=2.5, fig.align="center"}
ggplot() +
  geom_sf(data = HUC6.Florida, color = "darkgray", alpha = 0.5)+
  geom_sf(data = waterfeatures, aes(fill = FEATURE, color = FEATURE)) +
  geom_sf(data = hyp2.spatial, alpha = 0.5) +
 # scale_color_viridis_c(option = "inferno") +
  labs(size = "Chemo staticity") +
  theme(legend.position = "right",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.margin = margin(0,0,0,0))
```


We created violin plots to visualize the ranges of discharge and nitrate in the different sites. We made individual violin plots for discharge and nitrate for each site, and noticed the difference between the canal plots and the other waterway types (Figure 2). The canal plots have values much more concentrated in their lower ranges, and contain less variation overall. Visualizing the differences in canal variable distributions alerts us that they may influence our later data analysis.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of discharge and nitrate values of a stream and a canal", fig.height=4, fig.width = 7}
SANTAviolin<- SANTAdata%>%
  select(Flow_Inst,Nitrate_mgl)%>%
  mutate(site = "SANTA")
SANTA_violin_dis<-ggplot(SANTAviolin)+
  geom_violin(aes(x=site,y=Flow_Inst))+
  labs(y=expression(paste("Discharge (ft"^"3","/s)")))
SANTA_violin_nit<-ggplot(SANTAviolin)+
  geom_violin(aes(x=site,y=Nitrate_mgl))+
  labs(y="Nitrate (mg/L)")

DRAINviolin<- DRAINdata%>%
  select(Flow_Inst,Nitrate_mgl)%>%
  mutate(site = "DRAIN")
DRAIN_violin_dis<-ggplot(DRAINviolin)+
  geom_violin(aes(x=site,y=Flow_Inst))+
  labs(y=expression(paste("Discharge (ft"^"3","/s)")))
DRAIN_violin_nit<-ggplot(DRAINviolin)+
  geom_violin(aes(x=site,y=Nitrate_mgl))+
  labs(y="Nitrate (mg/L)")

plot_grid(SANTA_violin_dis,SANTA_violin_nit, DRAIN_violin_dis,DRAIN_violin_nit, ncol = 2)
```

We then created combined violin plots of discharge and nitrate for all the sites. The CAL and SANTA sites had much more discharge than any other site so here we present the discharge plot without CAL and SANTA to better visualize the smaller discharges (Figure 3). A similar plot for nitrate was also made (Figure 4).

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of discharge values of all sites (except CAL and SANTA)", fig.height=3, fig.width = 8}
#discharge plot for all sites except two to visualize smaller discharges
ggplot(
  violin_comb%>%filter(site!="CAL", site!="SANTA"),aes(x=site,y=Flow_Inst))+
  geom_violin()+
  labs(y=expression(paste("Discharge (ft"^"3","/s)")))
```


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of nitrate concentration values of all sites", fig.height=3, fig.width = 8}
#discharge plot for all sites except two to visualize smaller discharges
ggplot(violin_comb,aes(x=site,y=Nitrate_mgl))+
  geom_violin()+
  labs(y="Nitrate (mg/L)")
```

Our final exploratory analysis consisted of plotting discharge and nitrate over time at all 21 of our sites. This served two purposes: one, it helped us get a preliminary sense of the diversity in distributions and behaviors at all of the sites, and two, it highlighted any possible gaps in the data that we would need to wrangle out. In general, we did not notice any trends between sites, indicating a deeper analysis was needed. 

This report includes four examples of the graphs we generated: one spring, one river, one canal, and one creek (Figure 5) - all to give a sense of the range of site types and the data trends at each.


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height=13, fig.width = 10, fig.cap = "Discharge and nitrate value patterns of example sites"}
#Spring
BLUEdata$dateTime <- as.Date(BLUEdata$dateTime)

BLUEplot <- ggplot(BLUEdata, aes(x= dateTime, y = Flow_Inst)) +
  geom_line(color = "blue") +
  labs(x = "", y = expression("Instantaneous Flow (ft"^3*"/s)"))

BLUENO3plot <- ggplot(BLUEdata, aes(x = dateTime, y = Nitrate_mgl)) +
  geom_line() +
  labs(x = "Date", y = "Instantaneous Nitrate (mg/l)")

BLUECombined <- plot_grid(BLUEplot, BLUENO3plot, ncol = 1)

BLUE.grob <- textGrob("Blue Springs Discharge/Nitrate Over Time", gp=gpar(fontface="bold", fontsize=15))


#River
CALdata$dateTime <- as.Date(CALdata$dateTime)

CALplot <- ggplot(CALdata, aes(x= dateTime, y = Flow_Inst)) +
  geom_line(color = "blue") +
  labs(x = "", y = expression("Instantaneous Flow (ft"^3*"/s)"))

CALNO3plot <- ggplot(CALdata, aes(x = dateTime, y = Nitrate_mgl)) +
  geom_line() +
  labs(x = "Date", y = "Instantaneous Nitrate (mg/l)")

CALCombined <- plot_grid(CALplot, CALNO3plot, ncol = 1)

CAL.grob <- textGrob("Caloosahatchee River Discharge/Nitrate Over Time",
                      gp=gpar(fontface="bold", fontsize=15))



#Canal
MAINdata$dateTime <- as.Date(MAINdata$dateTime)

MAINplot <- ggplot(MAINdata, aes(x= dateTime, y = Flow_Inst)) +
  geom_line(color = "blue") +
  labs(x = "", y = expression("Instantaneous Flow (ft"^3*"/s)"))

MAINNO3plot <- ggplot(MAINdata, aes(x = dateTime, y = Nitrate_mgl)) +
  geom_line() +
  labs(x = "Date", y = "Instantaneous Nitrate (mg/l)")

MAINCombined <- plot_grid(MAINplot, MAINNO3plot, ncol = 1)

MAIN.grob <- textGrob("Main Canal Discharge/Nitrate Over Time",
                     gp=gpar(fontface="bold", fontsize=15))


#Creek
TURKdata$dateTime <- as.Date(TURKdata$dateTime)

TURKplot <- ggplot(TURKdata, aes(x= dateTime, y = Flow_Inst)) +
  geom_line(color = "blue") +
  labs(x = "", y = expression("Instantaneous Flow (ft"^3*"/s)"))

TURKNO3plot <- ggplot(TURKdata, aes(x = dateTime, y = Nitrate_mgl)) +
  geom_line() +
  labs(x = "Date", y = "Instantaneous Nitrate (mg/l)")

TURKCombined <- plot_grid(TURKplot, TURKNO3plot, ncol = 1)

TURK.grob <- textGrob("Turkey Creek Discharge/Nitrate Over Time",
                     gp=gpar(fontface="bold", fontsize=15))


grid.arrange(arrangeGrob(BLUECombined, top = BLUE.grob),
             arrangeGrob(CALCombined, top = CAL.grob),
             arrangeGrob(MAINCombined, top = MAIN.grob),
             arrangeGrob(TURKCombined, top = TURK.grob))
```


\newpage

# Analysis

## Question 1: Are sites with high hydrologic flashiness more or less likely to have high chemical flashiness?

To analyze our first question, on the relationship between hydrolic flashiness and chemical flashiness, we calculated the Richards-Baker Index of each site to quantify hydrologic flashiness. We had to exclude catchment size in the formula because 17 out of 21 sites did not have a recorded catchment size.

To quantify chemical flashiness, we ran a linear regression between mean daily nitrate and mean daily discharge values and took the coefficient as our chemostaticity value. We excluded 3 sites with poor linear regression fits.

We wrangled the RBI values and Chemostaticity Coefficients to create a dataframe with columns site, site type (canal, river, etc.), chemostaticity coefficient, chemostaticity coefficient absolute value, and RBI value. We included the absolute value of the Chemostaticity Coefficient to investigate the overall flashiness, as absolute value would account for large chemostaticity in either a flushing or diluting system.

```{r, include = FALSE}
# read in data from RBI_CHEMO_analysis doc
RBI_CHEMO_data<-read_csv("../Processed_Dataframes/RBI_CHEMO_data.csv")
RBI_nocanals<-read_csv("../Processed_Dataframes/RBI_CHEMO_nocanals.csv")  
```

We created ggplot visualizations of Chemostatic Coefficient vs. RBI (Figure 6) and Absolute Chemostatic Coefficient vs. RBI (Figure 7) separating site type by color. The visualization without absolute value shows a distribution that follows the x and y axis. Sites with large RBI values have low Chemostatic Coefficients, and sites with low RBI values have Chemostatic Coefficients that are far above 0 or far below 0. The absolute value visualization reinforces this finding, as large absolute Chemostatic Coefficients correspond with small RBI values. Similar distributions were viewed between site type, except that the largest outlier for both Chemostatic Coefficient and RBI are canals.

```{r, echo = FALSE, message = FALSE, fig.cap="Chemostatic coefficient values and Richards-Baker Index of 18 sites", fig.height=2.5}

#visualize data, both abs and non-abs
ggplot(RBI_CHEMO_data, 
                   aes(x=RBI_value_ind_noncatch, y=chemo.coef))+
  geom_point(aes(color=site.type))+
  labs(x="Richards-Baker Index", y="Chemostatic Coefficient")+theme(plot.margin = margin(0,0,0,0))
```


```{r, echo = FALSE, message = FALSE, fig.cap="Absolute Chemostatic coefficient values and Richards-Baker Index of 18 sites",fig.height=2.5}
ggplot(RBI_CHEMO_data, 
                       aes(x=RBI_value_ind_noncatch, y=chemo.coef.abs))+
  geom_point(aes(color=site.type))+
  labs(x="Richards-Baker Index", 
       y=" Absolute Chemostatic Coefficient")+theme(plot.margin = margin(0,0,0,0))

```

We ran a linear regression on log(Absolute Chemostatic Coefficient) vs. RBI, which was the closest fit that we could find. The regression produced a coefficient of -2.6847 (16 degrees of freedom, P-value of 0.473, multiple r-squared 0.03264). The model does not adequately predict the data.

```{r, echo = FALSE, message = FALSE, include = FALSE}
# run regressions on data

lm_sig_log_abs<-lm(log(chemo.coef.abs)~RBI_value_ind_noncatch, RBI_CHEMO_data)

summary(lm_sig_log_abs)

lm_nocanals_log_abs<-
  lm(log(chemo.coef.abs)~RBI_value_ind_noncatch, RBI_nocanals)
summary(lm_nocanals_log_abs)
```

Since canals often act differently than natural waterways, and the largest outlier for both RBI and Chemostaticity Coefficient are canals, we also visualized the data without canals (Figure 8). The visualizations show a more concise version of the trends in Figure 6 and Figure 7, where low RBI results in a high Absolute Chemostatic Coefficient. We ran the same linear regression on the data without canals.

```{r, echo = FALSE, message = FALSE, fig.cap="Absolute Chemostatic coefficient values and Richards-Baker Index of 14 non-canal sites",fig.height=2.5}
#same analysis, without canals in the data
Nocanals.linear.abs<-ggplot(RBI_nocanals, 
                            aes(x=RBI_value_ind_noncatch, y=chemo.coef.abs))+
  geom_point(aes(color=site.type))+
  geom_smooth(method="lm", se=TRUE)+
  labs(x="Richards-Baker Index",
       y="Absolute Chemostatic Coefficient")

print(Nocanals.linear.abs)
```

The linear regression for Absolute Chemostatic Coefficient vs. RBI provided a Coefficient of -6.889 (12 degrees of freedom, P-value 0.39, and a multiple r-squared 0.06228). While the fit still isn't significant, it is better than the data with canals and suggests that there is a negative relationship between hydrologic flashiness and chemical flashiness.


## Question 2: How do features of a site impact its chemical flashiness?

### Qualitative Approach: How do geographical features impact the site’s chemical flashiness?

We made observations by first sorting sites into three categories - Chemostatic, Flushing, and Diluting - based on the slopes of the regression lines of their C-Q plots - values near 0, positive values, and negative values, respectively. Then, we went site by site and searched their coordinates into Google Maps and ArcGIS, and noted the general surrounding geography - is it mostly urban or rural? Is it in a state park, adjacent to it, or nowhere near? Where, generally, in Florida is this site located? We also paid special attention to the site type - river, spring, or canal - as each most likely have differing management strategies and hydrologies.

When we were done listing the qualitative traits of all the sites in each hydrologic behavioral categorical, we noticed some trends had emerged. For the most flushing sites, or, the ones with the steepest regression line slopes, we noticed that all are located near airports. (Main Canal, Fellsmere Canal, Crane Creek, and Madison Blue Spring) There does not seem to be a correlation with the urbanity of the location, however, as two out of four of these sites are mostly rural and two are mostly urban - Fellsmere, Madison Blue Spring and Main Canal, Crane Creek, respectively. We hesitate to draw any concrete conclusions about possible relationships between flushing systems and proximity to airports, as there may be many other factors involved that are not within the scope of this project. One important factor could be the management style of the site; three out of four of the canals analyzed in this project tend towards flushing. This, too, could be a coincidence, and we would need to do a deeper analysis with more sites to draw conclusions.

Our list of diluting sites show more convincing trends - seven out of nine of the sites are located in very rural locations, most in or on the edge of a Florida state park. (Ichetucknee, Santa Fe, Blue Spring, Blue Hole Spring, Manatee Spring, Homosassa Spring, Turkey Creek) However, this number may be slightly inflated because three of the four most diluting sites are directly interconnected, increasing the likelihood of similar hydrologic tendencies between them. (Ichetucknee, Santa Fe, Blue Hole Spring) There seem to be no other notable qualitative trends for diluting sites.

Chemostatic or mostly chemostatic systems do not seem to display any particular kind of trend. They are located in all kinds of environments - from hyper-urban to deep in state parks.

```{r, include = FALSE, warning= FALSE, message = FALSE, echo = FALSE}
#Blue
BLUENitrateDV <- BLUEdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

BLUEFlowDV <- BLUEdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

BLUEChemo <- full_join(BLUENitrateDV, BLUEFlowDV, by = "dateTime")

BLUEChemoplot <- ggplot(BLUEChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Blue Springs") +
  theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#BOW
BOWNitrateDV <- BOWdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

BOWFlowDV <- BOWdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

BOWChemo <- full_join(BOWNitrateDV, BOWFlowDV, by = "dateTime")

BOWChemoplot <- ggplot(BOWChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  
  labs(x = "", y = "") +
  ggtitle("Rainbow River \n Near Dunnellon")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#CAL
CALNitrateDV <- CALdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

CALFlowDV <- CALdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

CALChemo <- full_join(CALNitrateDV, CALFlowDV, by = "dateTime")

CALChemoplot <- ggplot(CALChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Caloosahatchee River")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#CHAS
CHASNitrateDV <- CHASdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

CHASFlowDV <- CHASdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

CHASChemo <- full_join(CHASNitrateDV, CHASFlowDV, by = "dateTime")

CHASChemoplot <- ggplot(CHASChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Chassahowitzka River")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#CRANE
CRANENitrateDV <- CRANEdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

CRANEFlowDV <- CRANEdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

CRANEChemo <- full_join(CRANENitrateDV, CRANEFlowDV, by = "dateTime")

CRANEChemoplot <- ggplot(CRANEChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Crane Creek")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))
#DRAIN
DRAINNitrateDV <- DRAINdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

DRAINFlowDV <- DRAINdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

DRAINChemo <- full_join(DRAINNitrateDV, DRAINFlowDV, by = "dateTime")

DRAINChemoplot <- ggplot(DRAINChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Drainage Canal")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))  

#FAN
FANNitrateDV <- FANdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

FANFlowDV <- FANdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

FANChemo <- full_join(FANNitrateDV, FANFlowDV, by = "dateTime")

FANChemoplot <- ggplot(FANChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Fanning Springs")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#FELL
FELLNitrateDV <- FELLdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

FELLFlowDV <- FELLdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

FELLChemo <- full_join(FELLNitrateDV, FELLFlowDV, by = "dateTime")

FELLChemoplot <- ggplot(FELLChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Fellsmere Canal")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#HOL
HOLNitrateDV <- HOLdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

HOLFlowDV <- HOLdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

HOLChemo <- full_join(HOLNitrateDV, HOLFlowDV, by = "dateTime")

HOLChemoplot <- ggplot(HOLChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Blue Hole Spring")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#HOM
HOMNitrateDV <- HOMdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

HOMFlowDV <- HOMdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

HOMChemo <- full_join(HOMNitrateDV, HOMFlowDV, by = "dateTime")

HOMChemoplot <- ggplot(HOMChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Homosassa Springs")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#HUN
HUNNitrateDV <- HUNdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

HUNFlowDV <- HUNdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

HUNChemo <- full_join(HUNNitrateDV, HUNFlowDV, by = "dateTime")

HUNChemoplot <- ggplot(HUNChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Hunter Spring Run")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#ICHE
ICHENitrateDV <- ICHEdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

ICHEFlowDV <- ICHEdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

ICHEChemo <- full_join(ICHENitrateDV, ICHEFlowDV, by = "dateTime")

ICHEChemoplot <- ggplot(ICHEChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Ichetucknee River")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#MAD
MADNitrateDV <- MADdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

MADFlowDV <- MADdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

MADChemo <- full_join(MADNitrateDV, MADFlowDV, by = "dateTime")

MADChemoplot <- ggplot(MADChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Madison Blue Spring")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#MAIN
MAINNitrateDV <- MAINdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

MAINFlowDV <- MAINdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

MAINChemo <- full_join(MAINNitrateDV, MAINFlowDV, by = "dateTime")

MAINChemoplot <- ggplot(MAINChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Main Canal")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#MAN

MANNitrateDV <- MANdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

MANFlowDV <- MANdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

MANChemo <- full_join(MANNitrateDV, MANFlowDV, by = "dateTime")

MANChemoplot <- ggplot(MANChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Manatee Spring")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#RAIN
RAINNitrateDV <- RAINdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

RAINFlowDV <- RAINdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

RAINChemo <- full_join(RAINNitrateDV, RAINFlowDV, by = "dateTime")

RAINChemoplot <- ggplot(RAINChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  
  labs(x = "", y = "") +
  ggtitle("Rainbow River at Dunnellon")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#SANTA
SANTANitrateDV <- SANTAdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

SANTAFlowDV <- SANTAdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

SANTAChemo <- full_join(SANTANitrateDV, SANTAFlowDV, by = "dateTime")

SANTAChemoplot <- ggplot(SANTAChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Santa Fe River")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))


#THREE
THREENitrateDV <- THREEdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

THREEFlowDV <- THREEdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

THREEChemo <- full_join(THREENitrateDV, THREEFlowDV, by = "dateTime")

THREEChemoplot <- ggplot(THREEChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Three Mile Canal")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#TURK
TURKNitrateDV <- TURKdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

TURKFlowDV <- TURKdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

TURKChemo <- full_join(TURKNitrateDV, TURKFlowDV, by = "dateTime")

TURKChemoplot <- ggplot(TURKChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Turkey Creek")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

#TURN
TURNNitrateDV <- TURNdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

TURNFlowDV <- TURNdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

TURNChemo <- full_join(TURNNitrateDV, TURNFlowDV, by = "dateTime")

TURNChemoplot <- ggplot(TURNChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Turnbull Creek")+
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))
#WAC
WACNitrateDV <- WACdata %>%
  group_by(dateTime) %>%
  summarise(Nitrate_DV = mean(Nitrate_mgl))

WACFlowDV <- WACdata %>%
  group_by(dateTime) %>%
  summarise(Flow_DV = mean(Flow_Inst))

WACChemo <- full_join(WACNitrateDV, WACFlowDV, by = "dateTime")

WACChemoplot <- ggplot(WACChemo, aes(x = Flow_DV, y = Nitrate_DV)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "", y = "") +
  ggtitle("Wacissa River") +
theme(plot.title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

y.grob <- textGrob("Daily Mean Nitrate (mg/l)", 
                   gp=gpar(fontface="bold", fontsize=10), rot=90)

x.grob <- textGrob(("Daily Mean Discharge (ft3/s)"), 
                   gp=gpar(fontface="bold", fontsize=10))

diluting.grob <- textGrob("Diluting Systems C-Q",
                          gp=gpar(fontface="bold", fontsize=12))
flushing.grob <- textGrob("Flushing Systems C-Q",
                          gp=gpar(fontface="bold", fontsize=12))
chemostatic.grob <- textGrob("Chemostatic Systems C-Q",
                             gp=gpar(fontface="bold", fontsize=12))
```

```{r, echo = FALSE, message = FALSE,warning = FALSE, fig.cap="C-Q Plots of Chemostatic Sites",fig.height=3.2}

ChemostaticPlot <- plot_grid(
  FANChemoplot, RAINChemoplot, CHASChemoplot,  
  TURNChemoplot, WACChemoplot,
  ncol = 3)
grid.arrange(arrangeGrob(ChemostaticPlot, left = y.grob, bottom = x.grob, top = chemostatic.grob))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap="C-Q Plots of Flushing Sites",fig.height=4.2}
FlushingPlot <- plot_grid(
  MAINChemoplot, FELLChemoplot, CRANEChemoplot, MADChemoplot, 
  THREEChemoplot, CALChemoplot, HUNChemoplot,
  ncol = 3
)
grid.arrange(arrangeGrob(FlushingPlot, left = y.grob, bottom = x.grob, top = flushing.grob))

```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap="C-Q Plots of Diluting Sites",fig.height=4.2}
DilutingPlot <- plot_grid(
  ICHEChemoplot, SANTAChemoplot, BLUEChemoplot, HOLChemoplot, DRAINChemoplot, 
  BOWChemoplot,  MANChemoplot, HOMChemoplot, TURKChemoplot,
  ncol = 3
)
grid.arrange(arrangeGrob(DilutingPlot, left = y.grob, bottom = x.grob, top = diluting.grob))

```

### Quantitative/Statistical Approach: How do counties’ population and water usage impact the chemical flashiness of sites within the county?

Figure 12 shows each site's chemostaticity (absolute chemostatic coefficient) on the water feature map. Based on the map, it appears that sites on the right bank are less chemostatic, compared to sites in central-northern part of Florida. However, since many factors are confounding and sample size is not big enough, no definite conclusion can be drawn.

```{r, echo = FALSE, message = FALSE, fig.height=3.5, fig.width = 9,fig.cap="Location of sites and respective absolute chemostatic coefficient"}
hyp2.plot.2 <- ggplot() +
  geom_sf(data = HUC6.Florida, color = "darkgray", alpha = 0.5)+
  geom_sf(data = waterfeatures, aes(fill = FEATURE, color = FEATURE, show.legend = "fill")) +
  geom_sf(data = hyp2.spatial, aes(size = abschemo), alpha = 0.5, show.legend = "point") +
 # scale_color_viridis_c(option = "inferno") +
  labs(size = "Absolute Chemostaticity") +
  theme(legend.position = "right",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8),
        legend.margin = margin(0,0,0,0))
print(hyp2.plot.2)
```

We then used county population and water use data from USGS to conduct a quantitative analysis on how sites in different locations have different levels of chemical flashiness. Specifically, for water use, we looked at billion gallons of water used per day for irrigation and thermo electricity generation in each county. We believe that the more water used for irrigation in a county, the higher probability the sites in that county have of being influenced by nutrient run-off from farms within the county. A large amount of water used for thermo electricity generation may either imply that the water is well-managed and therefore more chemically stable, or imply that pollution may come from the generation process and make the water chemically flashy.

Figure 13 shows that sites in counties with higher population seem to have a higher probability of having higher absolute chemostatic coefficient, which means that they are less chemically stable. However, relationship between water use and chemical flashiness is not convincing enough.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height=3.5, fig.width = 9, fig.cap = "Relationship between a site's chemical flashiness and population and water use of the county the site is in" }
pop.plot.2 = ggplot(hyp2.full_cleaned, aes(x = total_pop, y = abschemo)) +
  geom_point(aes(color = site.type)) +
  geom_smooth(method = "lm") +
  labs(x = "Population of the county (k people)", y = "Absolute Chemostatic Coefficient (slope from regression)")+
  theme(title = element_text(size = 8),
        plot.margin = margin(0,0,0,0))

irri.plot.2 = ggplot(hyp2.full_cleaned, aes(x = irrigation, y = abschemo)) +
  geom_point(aes(color = site.type)) +
  geom_smooth(method = "lm") +
  labs(x = "Irrigation water use (Bgal/d)", y = "")+
    theme(title = element_text(size = 8),
        plot.margin = margin(0,0,0,0),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


therm.plot.2 = ggplot(hyp2.full_cleaned, aes(x = thermoelectric, y = abschemo)) +
  geom_point(aes(color = site.type)) +
  geom_smooth(method = "lm") +
  labs(x = "Power gen. water use (Bgal/d)", y = "")+
    theme(title = element_text(size = 8),
        plot.margin = margin(0,0,0,0),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
legend <- get_legend(
  # create some space to the left of the legend
  pop.plot.2 + theme(legend.box.margin = margin(0, 0, 0, 0))
)

title <- ggdraw() + 
  draw_label(
    "Impact of Population and Water Use on Abs. Chemostatic Coeff.",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 0)
  )
plot_row <- plot_grid(pop.plot.2 + theme(legend.position="none"),
          irri.plot.2+ theme(legend.position="none"),
          therm.plot.2+ theme(legend.position="none"), legend, ncol = 4)

plot_grid(plot_row, ncol = 1,rel_heights = c(0.1, 1))
```

We ran a linear regression to quantify the relationship between population, water use and chemical flashiness. The regression result shows that the impact of water use (both irrigation and thermoelectric) on chemical flashiness is not significant. But for population, we got a coefficient of 1.057e-05, and a P-value of 0.06. That is, we are 94% confident that on average, adding one thousand people to a county, the absolute value of chemostatic coefficient will increase by 1. 

\newpage

# Summary and Conclusions
This examination of hydrologic and chemical flashiness on 21 water bodies in Florida yielded a few notable results. Correspondingly, there are a few recommendations that managers of rivers, springs, creeks, canals, etc. in Florida might want to consider. 

First, there is no clear or statistically significant correlation between hydrologic and chemical flashiness. Our research suggests that there could be a negative correlation between hydrologic flashiness and chemical flashiness, but further research is needed to develop a clear relationship. As a result, in order to maintain the health of their rivers and streams, it is recommended that river and stream managers start tracking or continue to track proxies for both hydrologic and chemical flashiness and not simply just for one or the other. For hydrologic flashiness, river and stream managers should track high frequency discharge levels during storms. For chemical flashiness, river and stream managers should track high frequency nitrate and conductance levels during storms. If further research confirmed a negative relationship between hydrologic flashiness and chemical flashiness, managers could better predict problems for certain streams and respond accordingly.

A second conclusion of this examination is that there exists a relationship between the location of a water body and its chemical flashiness behavior. More specifically, the water bodies that were located closest to airports tended to exhibit flushing behavior, while the water bodies located in or near state parks tended to exhibit diluting behavior. Also water bodies located in counties that have higher population tended to exhibit more chemical flashiness. However, due to the limited scope both in site number and variable number of this study, these trends may be misleading or misrepresentative of other processes at work. For Florida water managers, it is recommended to at least pay closer attention to the possible interactions between airports or similar, large human operations and watersheds - the diluting nature of systems in state parks may be important as well, but they are likely already closely monitored. 


\newpage

# References

(1) “Hydrologic Flashiness.” Kansas River Inventory. Accessed November 21, 2019. Web: http://www.kansasriverinventory.org/home/hydrologic-flashiness

(2) Creed, I.F., et al. (2015). “The river as a chemostat: fresh perspectives on dissolved organic matter flowing down the river continuum.” Can. J. Fish. Aquat. Sci. 72: 1272-1285.

(3) “Calculating the Cost of Weather and Climate Disasters.” NOAA: National Centers for Environmental Information. Accessed November 21, 2019. Web: https://www.ncei.noaa.gov/news/calculating-cost-weather-and-climate-disasters

(4) Smith, V. H. and Schindler, D. W. (2009). “Eutrophication science: where do we go from here?” Trends in Ecology and Evolution. 24, 201-207.

(5) Stephenson, C. “Addressing Eutrophication in Florida, one watershed at a time.” University of Florida: IFAS Extension. July 23, 2018. Accessed November 21, 2019. Web: https://nwdistrict.ifas.ufl.edu/nat/2018/07/23/addressing-eutrophication-in-florida-one-watershed-at-a-time/.

(6) “Chassahowitzka River and Coastal Swamps.” Southwest Florida Management District. Accessed November 21, 2019. Web: https://www.swfwmd.state.fl.us/recreation/chassahowitzka-river-and-coastal-swamps

(7) Chislock, M. F., Doster, E., Zitomer, R. A. & Wilson, A. E. (2013). “Eutrophication: Causes, Consequences, and Controls in Aquatic Ecosystems.” Nature Education Knowledge 4(4):10.

(8) Walker, S. “In World That Says It’s Cutting Nutrient Pollution, Progress is Lacking.” World Resources Institute. March 4, 2019. Accessed November 21, 2019. Web: https://www.wri.org/blog/2019/03/world-says-its-cutting-nutrient-pollution-progress-lacking

(9) “Extreme weather gets a boost from climate change.” Environmental Defense Fund. Accessed November 21, 2019. Web: https://www.edf.org/climate/climate-change-and-extreme-weather

(10) Begley, J. “Pioneers had foresight to drain area for development; set up canal system in use today.” TC Palm News. November 7, 2018. Accessed November 21, 2019. Web: https://www.tcpalm.com/story/news/local/verobeachcentennial/2018/11/07/without-canals-dug-vero-beach-area-1900-s-wed-under-water/1821047002/

(11) “City of NSB Voter Information: Turnbull Creek.” Marine Discovery Center. October 10, 2018. Accessed November 21, 2019. Web: https://www.marinediscoverycenter.org/city-of-nsb-voter-information-turnbull-creek/

(12) Harrison, C. “NSB responds to growth with effort to preserve Turnbull Creek.” June 14, 2018. Accessed November 21, 2019. The Daytona Beach News Journal. Web:
https://www.news-journalonline.com/news/20180613/nsb-responds-to-growth-with-effort-to-preserve-turnbull-creek

(13) Coston, D. “NSB To Purchase More Property Along Turnbull Creek.” WNDB: News Daytona Beach. June 27, 2019. Accessed November 21, 2019. Web: https://www.newsdaytonabeach.com/wndb-news/nsb-to-purchase-more-property-along-turnbull-creek/ 

(14) Division of Environmental Assessment and Restoration, Water Quality Restoration Program, Department of Florida. “Basin Management Action Plan for the Implementation of the Nutrient Total Maximum Daily Load by the Florida Department of Environmental Protection in the Jackson Blue Spring and Merritts Mill Pond Basin.” May 2016. Accessed November 21, 2019. Web. https://floridadep.gov/sites/default/files/JacksonBlueBMAP.pdf

(15) Florida Department of Environmental Protection. “Protecting Florida’s Springs.” Accessed November 21, 2019. Web. https://floridadep.gov/springs/protect-restore/content/protecting-floridas-springs


